Modal Verification Patterns for Systems

### 2025-06-23 paper.tex
+

}\todo{inline,color=pink}{I am well-aware what counts as as (including self-plagiarism). These were annotated with todos including cautions for changing the text.}

@@

On the other hand, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back using an undo journal that records the original contents before modification as abstacted in Figure \ref{fig:journal} or completed using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place. In other words, unlike our ability to distinguish all resources (disk-pages and diskblocks) either being consistent under a snapshot or not in filesystems with \textsf{CoW}, with in-place updates, journalling mechanism does not allow such a claim. In other words, as we see in Figure \ref{fig:journal}, the middle level disk-layout is composed of the updates to the disk

Unlike Essentially, successfully writing the new superblock for a new snapshot is analagous to completing a valid journal entry, except it \emph{also} has the effect of updating the live filesystem data due to making the ``journal entries'' valid filesystem data and metadata blocks (an idea that goes back to early work on log-structured filesystems~\cite{Rosenblum1992LFS}).
In fact, we can think of the atomicity of the recovery from the crash state conceptually equal to the single block disk write persisting the change from $\textsf{Super}_0$ to $\textsf{Super}_1$ in Figure \ref{fig:cowindex}, therefore, considering \textsf{CoW} semantics, we do not need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.
\paragraph{Memory Reclamation in Client Code} In a typical configuration, allocating, deallocating and referencing memory locations for a client code leave the memory layout in a form of DAG (directed-asyclic-graph) with a unique root which resembles the first stack pointer referencing the allocated memory. To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cow_index}, from which a path is created for a stack pointer reference to the allocated memory. To reclaim the allocated memory location that is no longer being accessed, systems uses a well-known method called \emph{reference-counting} in which a new reference to the allocated memory increases the reference counter on it w.r.t a certain regime. 
\paragraph{Consistency of Reclaimed Memory} A sound memory reclamation is based on the conditions where the unique is not \emph{shared}, and there is no \emph{reachability} from the root to any the references pointing to the allocated memory. Once the requirements are satisfied then the memory location can be reclaimed \emph{soundly} -- i.e., without violating any access through a reference to it.
\paragraph{Buffering} Buffering is a technique used in variety of systems (filesystem page caching, write buffer in multi-processor architectures, translation-lookaside-buffers for translated addresses etc.) to delay the costly operation (committing file pages, flushing buffered writes from write-cache to memory etc.) making updates consistent.
\paragraph{Consistency of Reads of Buffered Writes} Consistency of reads of any value potentially buffered depends on the semantics of the operation making them \emph{observable}. A typical architecture such as Intel-x86 considers the \emph{flush} of buffered writes \emph{non-deterministically} -- i.e., at any time at any number of buffered writes. On the other hand, a typical filesystem can follow a certain policy to commit the in-memory file pages to the disk.

>>>>

}\todo[inline,color=pink]{I am well-aware what counts as plagiarism (including self-plagiarism). These were annotated with todos including cautions for changing the text.}

@@

On the other hand, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back using an undo journal that records the original contents before modification as abstacted in Figure \ref{fig:journal} or completed using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place. In other words, unlike our ability to distinguish all resources (disk-pages and dis-kblocks) either being consistent under a snapshot or not in filesystems with \textsf{CoW}, with in-place updates, journalling mechanism does not allow such a claim. In other words, as we see in Figure \ref{fig:journal}, unlike \textsf{CoW}'s uniform persistency claim over disk resources under a snapshot as a resource container, elaborative bookkeeping mechanism of journalling leaves the filesystem treat disk resources in one of the resource groups in non-uniform fashion: disk-pages loaded into memory, updated disk-page not logged yet, updated and being logged or already logged disk-pages. Being able to treat the resources in a uniform fashion determines wwhether we need to consider the specification and proof of crash condition separately; instead a global never-violated invariant (not even temporarily broken) of the filesystem ensures the on-disk state is always consistent.

\paragraph{Memory Reclamation in Client Code} In a typical configuration, allocating, deallocating and referencing memory locations for a client code leave the memory layout in a form of DAG (directed-asyclic-graph) with a unique root which resembles the first stack pointer referencing the allocated memory. To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cowindex}, from which a path is created for a stack pointer reference to the allocated memory. To reclaim the allocated memory location that is no longer being accessed, systems uses a well-known method called \emph{reference-counting} in which a new reference to the allocated memory increases the reference counter on it w.r.t a certain regime. 
\paragraph{Consistency of Reclaimed Memory} A sound memory reclamation is based on the conditions where the unique is not \emph{shared}, and there is no \emph{reachability} from the root to any the references pointing to the allocated memory. Once the requirements are satisfied then the memory location can be reclaimed \emph{soundly} -- i.e., without violating any access through a reference to it.
\paragraph{Buffering} Buffering is a technique used in variety of systems (filesystem page caching, write buffer in multi-processor architectures, translation-lookaside-buffers for translated addresses etc.) to delay the costly operation (committing file pages, flushing buffered writes from write-cache to memory etc.) making updates consistent.
\paragraph{Consistency of Reads of Buffered Writes} Consistency of reads of any value potentially buffered depends on the semantics of the operation making them \emph{observable}. A typical architecture such as Intel-x86 considers the \emph{flush} of buffered writes \emph{non-deterministically} -- i.e., at any time at any number of buffered writes. On the other hand, a typical filesystem can follow a certain policy to commit the in-memory file pages to the disk.
